### 2009-03-19 doc/storage_system/arc-storage-manual.tex
starting a quick start manual to install and configure the storage system

\tableofcontents                          %Comment if use article style
\begin{verbatim}
    Administrator's and user's manual
    ---------------------------------

    Administrator's manual
    - Quick start guide
        - Checking the code out from the subversion
        - Compiling and installing
        - Modifying the configuration template
        - Running the daemon
        - Using the CLI tool
    - Installation:
        - Getting the source code from subversion
        - Compilation details
        - Installation
    - Configuration:
        - HED service config
        - Centralized A-Hash
        - Distributed A-Hash
        - Librarian
        - Shepherd
        - Hopi
        - Bartender
        - Configuring inter-service trust: local listing of DNs, or using an A-Hash to store them
        - Virtual organizations: how to configure HED to trust a VO
    - Deployment scenarios:
        - Single node: putting every service on one single node (including a storage node with Hopi)
        - Centralized management: using seperate multiple storage nodes with Hopi, but leaving a central A-Hash, a Librarian and a Bartender on the central node
        - Simple distribution: split the central node to three nodes, each of them has an A-Hash, a Librarian and a Bartender, and we have separate storage nodes
        - One service per node: we have seperate nodes for all the services, any number of A-Hash nodes, any number of Librarian nodes, Bartender nodes, storage nodes

    User's manual
    - Overview: what do you need to access a running ARC storage system
    - Prototype command line tool
        - Configuring the command line tool
        - Creating the root collection
        - Uploading files
        - Creating collections
        - Moving files
        - Creating hardlinks
        - Removing files
        - Removing collections
        - Close collections
        - Modify access policies
        - Working with mount points
    - Grid jobs
        - Configuring the ARC storage DMC
        - Specifying files in job descriptions
    - Accessing the system with FUSE
        - Configuring the FUSE module
        - Mounting the storage system
        - Supported operations
        - Limitations

@@

In this section we will follow the installation of the ARC storage system on a freshly installed debian linux machine.

@@

\begin{verbatim}
$ cd arc1
$ ./autogen.sh
$ ./configure --disable-java --disable-a-rex-service --disable-isi-service --disable-charon-service --disable-compiler-service --disable-paul-service --disable-sched-service
--prefix

@@

@@

\section{Installation} % (fold)
\label{sec:installation}

% section installation (end)

\section{Configuration} % (fold)
\label{sec:configuration}

% section configuration (end)

\section{Deployment scenarios} % (fold)
\label{sec:deployment_scenarios}

% section deployment_scenarios (end)

@@

\section{Overview} % (fold)
\label{sec:overview}

% section overview (end)

\section{Prototype command line tool} % (fold)
\label{sec:prototype_command_line_tool}

% section prototype_command_line_tool (end)

\section{Grid jobs} % (fold)
\label{sec:grid_jobs}

% section grid_jobs (end)

\section{Accessing the system with FUSE} % (fold)
\label{sec:accessing_the_system_with_fuse}

% section accessing_the_system_with_fuse (end)

>>>>

\tableofcontents                          %Comment if use article style
% \begin{verbatim}
%     Administrator's and user's manual
%     ---------------------------------
% 
%     Administrator's manual
%     - Quick start guide
%         - Checking the code out from the subversion
%         - Compiling and installing
%         - Modifying the configuration template
%         - Running the daemon
%         - Using the CLI tool
%     - Installation:
%         - Getting the source code from subversion
%         - Compilation details
%         - Installation
%     - Configuration:
%         - HED service config
%         - Centralized A-Hash
%         - Distributed A-Hash
%         - Librarian
%         - Shepherd
%         - Hopi
%         - Bartender
%         - Configuring inter-service trust: local listing of DNs, or using an A-Hash to store them
%         - Virtual organizations: how to configure HED to trust a VO
%     - Deployment scenarios:
%         - Single node: putting every service on one single node (including a storage node with Hopi)
%         - Centralized management: using seperate multiple storage nodes with Hopi, but leaving a central A-Hash, a Librarian and a Bartender on the central node
%         - Simple distribution: split the central node to three nodes, each of them has an A-Hash, a Librarian and a Bartender, and we have separate storage nodes
%         - One service per node: we have seperate nodes for all the services, any number of A-Hash nodes, any number of Librarian nodes, Bartender nodes, storage nodes
% 
%     User's manual
%     - Overview: what do you need to access a running ARC storage system
%     - Prototype command line tool
%         - Configuring the command line tool
%         - Creating the root collection
%         - Uploading files
%         - Creating collections
%         - Moving files
%         - Creating hardlinks
%         - Removing files
%         - Removing collections
%         - Close collections
%         - Modify access policies
%         - Working with mount points
%     - Grid jobs
%         - Configuring the ARC storage DMC
%         - Specifying files in job descriptions
%     - Accessing the system with FUSE
%         - Configuring the FUSE module
%         - Mounting the storage system
%         - Supported operations
%         - Limitations
%     
% \end{verbatim}

@@

In this section we will follow the installation of the ARC storage system on a freshly installed debian lenny machine.

@@

\begin{verbatim}
$ cd arc1
$ ./autogen.sh
$ ./configure --disable-java --disable-a-rex-service --disable-isi-service \
    --disable-charon-service --disable-compiler-service \
    --disable-paul-service --disable-sched-service

@@

It is important to syncronize the time of the machine - the storage system will make decisions based on differences between timestamps, when we deploy it on multiple machines, different times would cause problems.

We need host (and user) certificates to run the system. For testing purposes we can use the NorduGrid InstantCA solution which is capable of creating a demo-CA with short lifetime. The URL of the InstantCA service is \verb!https://vls.grid.upjs.sk/CA/instantCA!, let's create 3 user and 3 host certificates. (e.g. we can set the Organization Name to `knowarc', the Common name of the CA to `storage-test', the name of the users could be `penny', `billy' and `hammer', the name of the hosts could be `upsilon', `omicron' and `theta', and we need a password for the CA and passwords for the user certificates, both should be at least 4 characters long).

Let's download the generated certificates and untar the archive file to see its contents:

\begin{verbatim}
$ ls -lR
.:
total 28
drwxr-xr-x 2 zsombor zsombor 4096 2009-03-19 12:11 CA
-rw-r--r-- 1 zsombor zsombor  963 2009-03-19 12:11 ca.key
-rw-r--r-- 1 zsombor zsombor  786 2009-03-19 12:11 ca.pem
drwxr-xr-x 2 zsombor zsombor 4096 2009-03-19 12:11 hostCerts
-rw-r--r-- 1 zsombor zsombor 1044 2009-03-19 12:11 readme
-rw-r--r-- 1 zsombor zsombor    3 2009-03-19 12:11 serial.srl
drwxr-xr-x 2 zsombor zsombor 4096 2009-03-19 12:11 userCerts

./CA:
total 8
-rw-r--r-- 1 zsombor zsombor 786 2009-03-19 12:11 66fe707c.0
-rw-r--r-- 1 zsombor zsombor 139 2009-03-19 12:11 66fe707c.signing_policy

./hostCerts:
total 24
-rw-r--r-- 1 zsombor zsombor 786 2009-03-19 12:11 hostcert-omicron.pem
-rw-r--r-- 1 zsombor zsombor 782 2009-03-19 12:11 hostcert-theta.pem
-rw-r--r-- 1 zsombor zsombor 786 2009-03-19 12:11 hostcert-upsilon.pem
-rw-r--r-- 1 zsombor zsombor 887 2009-03-19 12:11 hostkey-omicron.pem
-rw-r--r-- 1 zsombor zsombor 891 2009-03-19 12:11 hostkey-theta.pem
-rw-r--r-- 1 zsombor zsombor 887 2009-03-19 12:11 hostkey-upsilon.pem

./userCerts:
total 24
-rw-r--r-- 1 zsombor zsombor 778 2009-03-19 12:11 usercert-billy.pem
-rw-r--r-- 1 zsombor zsombor 778 2009-03-19 12:11 usercert-hammer.pem
-rw-r--r-- 1 zsombor zsombor 778 2009-03-19 12:11 usercert-penny.pem
-rw-r--r-- 1 zsombor zsombor 963 2009-03-19 12:11 userkey-billy.pem
-rw-r--r-- 1 zsombor zsombor 963 2009-03-19 12:11 userkey-hammer.pem
-rw-r--r-- 1 zsombor zsombor 951 2009-03-19 12:11 userkey-penny.pem
\end{verbatim}

We have certificate and key files for all the hosts and users, and we have the CA file with the proper hashed name.
There is no rule where to put these certificate, but now we will put them in \verb!/etc/grid-security!. Let's create this directory, and a subdirectory called \verb!certificates!. We will use the `theta' host certificates for this machine, let's copy it there:

\begin{verbatim}
$ sudo cp hostCerts/hostcert-theta.pem /etc/grid-security/hostcert.pem
$ sudo cp hostCerts/hostkey-theta.pem /etc/grid-security/hostkey.pem
$ sudo cp CA/* /etc/grid-security/certificates/
$ ls -lR /etc/grid-security/
/etc/grid-security/:
total 12
drwxr-xr-x 2 root root 4096 2009-03-19 12:36 certificates
-rw-r--r-- 1 root root  782 2009-03-19 12:35 hostcert.pem
-rw-r--r-- 1 root root  891 2009-03-19 12:36 hostkey.pem

/etc/grid-security/certificates:
total 8
-rw-r--r-- 1 root root 786 2009-03-19 12:36 66fe707c.0
-rw-r--r-- 1 root root 139 2009-03-19 12:36 66fe707c.signing_policy
\end{verbatim}

And I choose a user certificate and put it into the \verb!~/.arc! directory, conveniently removing the password from the key file:

\begin{verbatim}
$ mkdir ~/.arc
$ cp userCerts/usercert-billy.pem ~/.arc/usercert.pem
$ openssl rsa -in userCerts/userkey-billy.pem -out ~/.arc/userkey.pem
$ chmod 600 ~/.arc/userkey.pem
$ ls -l ~/.arc
total 8
-rw-r--r-- 1 zsombor zsombor 778 2009-03-19 13:06 usercert.pem
-rw------- 1 zsombor zsombor 891 2009-03-19 13:09 userkey.pem
\end{verbatim}

The storage system runs withen the \verb!arched! hosting environment daemon, which needs a configuration file which describes which services we want to run, and on which ports do we want to listen, etc.
Let's copy the template configuration files to the \verb!/etc/arc! directory:

\begin{verbatim}
$ sudo mkdir /etc/arc
$ sudo cp arc1/src/services/storage/storage_service.xml.example /etc/arc/storage_service.xml
$ sudo cp arc1/src/services/storage/clientsslconfig.xml /etc/arc
$ ls -l /etc/arc
total 12
-rw-r--r-- 1 root root  259 2009-03-19 13:26 clientsslconfig.xml
-rw-r--r-- 1 root root 4606 2009-03-19 13:25 storage_service.xml
\end{verbatim}

The template configuration file specifies several directories where the services can store their data. We will create these directories at \verb!/var/spool/arc!, but they can be created anywhere else.

\begin{verbatim}
$ sudo mkdir /var/spool/arc/ahash_data
$ sudo mkdir /var/spool/arc/shepherd_data
$ sudo mkdir /var/spool/arc/shepherd_store
$ sudo mkdir /var/spool/arc/shepherd_transfer
$ ls -l /var/spool/arc
total 16
drwxr-xr-x 2 root root 4096 2009-03-19 13:42 ahash_data
drwxr-xr-x 2 root root 4096 2009-03-19 13:42 shepherd_data
drwxr-xr-x 2 root root 4096 2009-03-19 13:42 shepherd_store
drwxr-xr-x 2 root root 4096 2009-03-19 13:42 shepherd_transfer
\end{verbatim}

In this deployment we plan to run the daemon as root, so it is OK that the root owns these directories. 
If we create these directories somewhere else, we would need to modify the paths in the \verb!storage_service.xml! file.

If we put the host certificates and the CA certificate somewhere else than \verb!/etc/grid-security! then we would need to modify the \verb!storage_service.xml! (look for \verb!KeyPath!, \verb!CertificatePath! and \verb!CACertificatesDir!), and we would need to modify \verb!/etc/arc/clientsslconfig.xml! as well. This file contains the credentials which is used by the services when they connect to an other service. If we put \verb!clientsslconfig.xml! somewhere else than \verb!/etc/arc! then we would need to modify all the \verb!ClientSSLConfig! tags in the \verb!storage_service.xml!.

Let's run the \verb!arched! daemon with this config (specified by the \verb!-c! option) first in the foreground (using the \verb!-f! option) to see immediately if everything is right:

\begin{verbatim}
$ sudo /usr/local/sbin/arched -c /etc/arc/storage_service.xml -f
\end{verbatim}

The template configuration file specifies a loglevel which only prints error messages, so nothing should be written on the screen now.
The \verb!arched! daemon should now listen on the ports: 60001 for our data transfer service called Hopi, and 60000 for all the other services:

\begin{verbatim}
$ netstat -at | grep LISTEN
tcp        0      0 *:60000                 *:*                     LISTEN     
tcp        0      0 *:60001                 *:*                     LISTEN     
\end{verbatim}

Now it is time to set up or prototype client tool to access the system. It is currently called \verb!arc_storage_cli! and it is by default installed to \verb!/usr/local/bin!. We have to tell this CLI tool where it can find our user credentials, and a URL of a Bartender service. The Bartender service is the front-end of the storage system. We can specify these things by environment variables, or by a configuration file called \verb!~/.arc/client.xml!. Let's create this \verb!client.xml! file:

\begin{verbatim}
$ cat ~/.arc/client.xml 
<ArcConfig>
  <KeyPath>~/.arc/userkey.pem</KeyPath>
  <CertificatePath>~/.arc/usercert.pem</CertificatePath>
  <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
  <BartenderURL>https://localhost:60000/Bartender</BartenderURL>
</ArcConfig>
\end{verbatim}

Now we can use the \verb!arc_storage_cli! to access our local deployment. Let's do a `list' on the root collection (`/'):

\begin{verbatim}
$ arc_storage_cli
Usage:
  arc_storage_cli <method> [<arguments>]
Supported methods: stat, make[Collection], unmake[Collection], list, move, put[File], get[File], del[File], pol[icy], unlink, credentialsDelegation, removeCredentials
Without arguments, each method prints its own help.

$ arc_storage_cli list
- The URL of the Bartender: https://localhost:60000/Bartender
- The key file: ~/.arc/userkey.pem
- The cert file: ~/.arc/usercert.pem
- The CA dir: /etc/grid-security/certificates
Usage: list <LN> [<LN> ...]

$ arc_storage_cli list /
- The URL of the Bartender: https://localhost:60000/Bartender
- The key file: ~/.arc/userkey.pem
- The cert file: ~/.arc/usercert.pem
- The CA dir: /etc/grid-security/certificates
- Calling the Bartender's list method...
[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544] Component tcp.client(tcp) could not be created
[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544] Component tls.client(tls) could not be created
[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544] Component http.client(http) could not be created
[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544] Component soap.client(soap) could not be created
ERROR: Wrong status from server.
Maybe wrong URL: 'https://localhost:60000/Bartender'

@@

% \section{Installation} % (fold)
% \label{sec:installation}
% 
% % section installation (end)
% 
% \section{Configuration} % (fold)
% \label{sec:configuration}
% 
% % section configuration (end)
% 
% \section{Deployment scenarios} % (fold)
% \label{sec:deployment_scenarios}
% 
% % section deployment_scenarios (end)

@@

% \section{Overview} % (fold)
% \label{sec:overview}
% 
% % section overview (end)
% 
% \section{Prototype command line tool} % (fold)
% \label{sec:prototype_command_line_tool}
% 
% % section prototype_command_line_tool (end)
% 
% \section{Grid jobs} % (fold)
% \label{sec:grid_jobs}
% 
% % section grid_jobs (end)
% 
% \section{Accessing the system with FUSE} % (fold)
% \label{sec:accessing_the_system_with_fuse}
% 
% % section accessing_the_system_with_fuse (end)
